Filling in data gaps on refugee children: Statistical imputation for missing count data

### 2023-01-13 paper.qmd
limitations chapter

---
title: "Filling in data gaps on refugee children: Statistical imputation for missing count data"
author: Jan Beise, Yukun Pei (UNICEF); Sebastian Steinmueller (UNHCR)
format: pdf
execute:
  echo: false
  warning: false
  message: false
header-includes:
    \usepackage{setspace}
    \onehalfspacing
---

@@

### read brms fit objects
m.ageonly.rds <- readRDS("models/m.ageonly.rds")
#m.age <- readRDS("models/m.age2_20230108.rds")

@@

# Calculate the proportions of children and girls/women
dem_longMissing <- dem_longMissing %>%
  mutate(childrenProp = children/total,
         femaleProp = female/total,
         missingAge = case_when(
              missing == "none" ~ "Age not missing",
              missing != "none" ~ "Age missing" 
            ),
         origin_countryShort = case_when(
              origin_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              origin_country == "United Kingdom of Great Britain and Northern Ireland" ~ "Bolivia",
              origin_country == "Bolivia (Plurinational State of)" ~ "UK",
              origin_country == "Central African Republic" ~ "CAR",
              origin_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              origin_country == "Lao People's Democratic Republic" ~ "Lao",
              origin_country == "Russian Federation" ~ "Russia",
              origin_country == "Syrian Arab Republic" ~ "Syria",
              origin_country == "Democratic Republic of the Congo" ~ "DRC",
              origin_country == "United Republic of Tanzania" ~ "Tanzania",
              origin_country == "United States of America" ~ "USA",
              origin_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(origin_country)
          ),
         asylum_countryShort = case_when(
              asylum_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              asylum_country == "United Kingdom of Great Britain and Northern Ireland" ~ "Bolivia",
              asylum_country == "Bolivia (Plurinational State of)" ~ "UK",
              asylum_country == "Central African Republic" ~ "CAR",
              asylum_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              asylum_country == "Lao People's Democratic Republic" ~ "Lao",
              asylum_country == "Russian Federation" ~ "Russia",
              asylum_country == "Syrian Arab Republic" ~ "Syria",
              asylum_country == "Democratic Republic of the Congo" ~ "DRC",
              asylum_country == "United Republic of Tanzania" ~ "Tanzania",
              asylum_country == "United States of America" ~ "USA",
              asylum_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(asylum_country)  
          )
         ) %>% 
  mutate(origin_countryShort = factor(origin_countryShort), 
         asylum_countryShort = factor(asylum_countryShort))

@@

## Demographic data and data gaps on refugees
UNHCR compiles official statistics on stocks and flows of forcibly displaced and stateless persons twice a year, once for mid-year figures (Mid-Year Statistical Reporting, MYSR) and once for end-year figures (Annual Statistical Reporting, ASR). For these reporting exercises, UNHCR's country offices compile aggregate population figures for refugees and asylum-seekers hosted in their country from a range of sources and data producers such as governments, UNHCR's own refugee registration database proGres and sometimes non-governmental actors. The geographic level on which theses statistics are being produced is therefore always the country country of asylum. The figures undergo a statistical quality control process at the country, regional and global level of the organisation and are disseminated on the publicly available refugee data finder (https://www.unhcr.org/refugee-statistics/) after applying statistical disclosure control to suppress very small counts of persons that could identify individuals. The models and results presented in this paper use data only on the end-year stock of refugees under UNHCR's mandate (including others in need of international protection and refugee-like populations^[For better readibility, we will refer to all populations used in the data as refugees.], excluding Palestine refugees under UNRWA's mandate), that is, not on asylum-seekers, internally displaced persons and stateless persons.

@@

@fig-missingAsy shows for what proportion of the refugee population living in each region age-disaggregated data was available at the end of 2021. While demographic coverage is close to universal for refugees hosted in Africa, it is available for `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Europe", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent of refugees in Europe, `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Asia", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent in Asia and only for `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Americas", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent in the Americas. This is to a large extent a result of the differing population data sources in these regions: While the individual demographic details of refugees in many countries in Africa and MENA are recorded in UNHCR’s own case registration system proGres, population data in other regions often comes from government offices with varying degrees of availability of demographic data.

@@

p.oriregion.missingAge2 <- ggplot(data = t.oriregion.missingAge %>%
                                   filter(origin_region!="Oceania", missingAge == "Age missing") ,
                                                 aes(x = origin_region, 
                                                     y = total)) +
  geom_bar(stat = "identity", colour = unhcrPaletteBlack[1], fill = unhcrPaletteWhite[1], alpha = 0,
           size = 1, linetype = "solid")  +
  ggtitle("Age data missing") +
 # labs(y = "Number of refugees")  +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45),
        axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  scale_y_continuous(labels = label_number(scale_cut = cut_si("m")))

@@

 p.top4.children.ori.asy
```
## Model
Considering the findings of the descriptive analysis with considerable differences in the distribution of countries of origin and asylum between refugees with and without available data, and the explanatory power of origins and countries of asylum for the demographic characteristics of refugee populations, ignoring the missing data (i.e. treating it as if it was the same as the available data) risks introducing bias into regional and global estimates of the proportion of refugee children. In the following, we propose a model to predict counts for the age/sex brackets of those country pairs with missing demographic data (the _NAs_ in @tbl-data), which in turn will allow us to estimate regional and global proportions of children. Summarising the descriptive analysis of missingness and available age data, there are three main conclusions to be considered in predicting the proportion of children for refugee populations without available age data:

1. We cannot assume the global proportion of children among refugees for whom demographic data is available is the same as the proportion we would see if data was available for populations with currently missing data.  
2. Knowing which country of origin refugees are from gives us some information regarding the proportion of children we would expect. For example, looking at @fig-childrenOriAsy we would generally assume that there is a higher proportion of children in a refugee population from South Sudan than in refugees from Venezuela.  
3. For each origin population, knowing which region of asylum a population is from can help us refine the estimate for the proportion of children based on the country of origin. If we know a population is from Syria and they reside in a country of asylum in Europe, we would assume the proportion of children to be lower than if they were living in a country in Asia.

@@

Multilevel models (also called hierarchical, or sometimes random effect, models) are well suited to pool information across geographic levels as described above. We fit a multilevel model with varying nested intercepts at the levels of country of origin, region of asylum and country of asylum. Since the data is available as aggregated counts in ten age/sex brackets that sum up to the population total, we choose a multinomial distribution to model the underlying probabilities of a refugee falling in each of the age/sex brackets depending on their country of origin and country of asylum. When estimating these probabilities, the model treats populations from the same origin country as being more similar to each other than those from different origins, and within each origin considers those residing in countries in the same region of asylum as more similar than those in different regions of asylum. We fit the model in a full Bayesian framework that allows us to incorporate uncertainty in parameters and in predictions from those parameters in the final predictions.

@@

The prior on the global intercepts with the $log(\frac{\Pi_{k}}{\Pi_{4}})$ mean aims to provide sensible levels for the age/sex bracket proportions. It incorporates our prior belief that, for example, it is unlikely that all refugees in a country are girls aged 0 to 4, but otherwise is not overly narrow.
The student-t priors on the group-level standard deviations of the intercepts of country of origin and region of asylum are not informative and only serve to regularise the parameter space. The prior on country of origin is slightly wider to allow less pooling (i.e. closer to a separate model for each country of origin) and thereby take into consideration the general idea of the model that age/sex distributions are idiosyncratic to populations of an origin country. 
The very narrow $N(0,0.01)$ prior on standard deviation of the country of asylum term $b_{k,a}$ merits further discussion: Including the term might seem unusual to start with since in most cases there is only one observation in the form of age/sex counts and total (i.e. one row in the dataset) per country of origin/asylum to start with (exception are countries in Latin America and the Caribbean with two populations displaced from Venezuela, refugees and others in need of international protection). On one hand, leaving out this parameter leads to overconfidence in predicting counts in country pairs with missing data since in that case the counts are drawn from the varying intercepts of the regions (which are observed in the data in  most country of origin/region of asylum pairs), ignoring the uncertainty that comes with predicting a whole new country pair not used in fitting the model. On the other hand, allowing a wider standard deviation leads to unreasonably wide prediction intervals (sometimes including counts of zero children in populations of several thousand refugees). Restricting the prior by way of a very small standard deviation helps us constrain the predictive intervals to sensible ranges.

To predict the count vector of age/sex brackets $\tilde{y_{i^{*}}}$ for a country pair $i^{*}$ without reported data and account for uncertainty of parameters $\theta$ and prediction, we draw for each country pair with missing data $m$ values from the posterior predictive distributions:
$$p(\tilde{y_{i^{*}}} \mid y)
= 
\int p(\tilde{y_{i^{*}}} \mid \theta) \cdot p(\theta \mid y)
\, \textrm{d}\theta$$

@@

The model was fit with the R-package brms. Reproducible code is available at https://github.com/sebastiansteinmueller/ref-demographics-imputation/.

## Results

```{r}
##### check predictions #####

imputations <- imputations %>% 
  ungroup() %>%
  mutate(femaleProp = female/total,
         children = rowSums(select(., female_0_4, female_5_11, female_12_17, male_0_4, male_5_11, male_12_17))) %>% 
  mutate(childrenProp = children/total)

@@

p.imputations.syr.children <- ggplot(data = imputations.syr %>% filter(asylum_iso3 %in% c("SWE", "USA",    "LBN", "JOR", "GBR")),
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_halfeye() +
  theme_minimal()

p.imputations.syr.reg.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_gradientinterval() +
  facet_wrap(~ asylum_region, scales = "free")

p.imputations.syr.neighbour.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_country, childrenProp), x=childrenProp)) +
     stat_gradientinterval() +
  facet_wrap(~ neighbor, scales = "free")

### by region of asylum
t.dem.asyreg.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  group_by(asylum_region) %>% 
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

@@

### world
t.dem.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

@@

## Limitations and outlook

>>>>

---
title: "Filling in data gaps on refugee children: Statistical imputation for missing count data"
author: Jan Beise, Yukun Pei (UNICEF); Sebastian Steinmueller (UNHCR)
format: pdf
execute:
  echo: false
  warning: false
  message: false
header-includes:
    \usepackage{float}
    \usepackage{setspace}
    \onehalfspacing
---

@@

### read brms fit object
m.ageonly <- readRDS("models/m.ageonly.rds")

@@

# Calculate the proportions of children and girls/women
dem_longMissing <- dem_longMissing %>%
  mutate(childrenProp = children/total,
         femaleProp = female/total,
         missingAge = case_when(
              missing == "none" ~ "Age not missing",
              missing != "none" ~ "Age missing" 
            ),
         origin_countryShort = case_when(
              origin_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              origin_country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
              origin_country == "Bolivia (Plurinational State of)" ~ "Bolivia",
              origin_country == "Central African Republic" ~ "CAR",
              origin_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              origin_country == "Lao People's Democratic Republic" ~ "Lao",
              origin_country == "Russian Federation" ~ "Russia",
              origin_country == "Syrian Arab Republic" ~ "Syria",
              origin_country == "Democratic Republic of the Congo" ~ "DRC",
              origin_country == "United Republic of Tanzania" ~ "Tanzania",
              origin_country == "United States of America" ~ "USA",
              origin_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(origin_country)
          ),
         asylum_countryShort = case_when(
              asylum_country == "Venezuela (Bolivarian Republic of)" ~ "Venezuela" ,
              asylum_country == "United Kingdom of Great Britain and Northern Ireland" ~ "UK",
              asylum_country == "Bolivia (Plurinational State of)" ~ "Bolivia",
              asylum_country == "Central African Republic" ~ "CAR",
              asylum_country == "China, Hong Kong Special Administrative Region" ~ "Hong Kong",
              asylum_country == "Lao People's Democratic Republic" ~ "Lao",
              asylum_country == "Russian Federation" ~ "Russia",
              asylum_country == "Syrian Arab Republic" ~ "Syria",
              asylum_country == "Democratic Republic of the Congo" ~ "DRC",
              asylum_country == "United Republic of Tanzania" ~ "Tanzania",
              asylum_country == "United States of America" ~ "USA",
              asylum_country == "United Arab Emirates" ~ "UAE",
              TRUE ~ as.character(asylum_country)  
          )
         ) %>% 
  mutate(origin_countryShort = factor(origin_countryShort), 
         asylum_countryShort = factor(asylum_countryShort))

@@

# Demographic data and data gaps on refugees

UNHCR compiles official statistics on stocks and flows of forcibly displaced and stateless persons twice a year, once for mid-year figures (Mid-Year Statistical Reporting, MYSR) and once for end-year figures (Annual Statistical Reporting, ASR). For these reporting exercises, UNHCR's country offices compile aggregate population figures for refugees and asylum-seekers hosted in their country from a range of sources and data producers such as governments, UNHCR's own refugee registration database proGres and sometimes non-governmental actors. The geographic level on which theses statistics are being produced is therefore always the country country of asylum. The figures undergo a statistical quality control process at the country, regional and global level of the organisation and are disseminated on the publicly available refugee data finder (https://www.unhcr.org/refugee-statistics/) after applying statistical disclosure control to suppress very small counts of persons that could identify individuals. The models and results presented in this paper use data only on the end-year stock of refugees under UNHCR's mandate (including others in need of international protection and refugee-like populations[^1], excluding Palestine refugees under UNRWA's mandate), that is, not on asylum-seekers, internally displaced persons and stateless persons.

[^1]: For better readability, we will refer to refugees, refugee-like populations and others in need of international protection as refugees throughout this text.

@@

@fig-missingAsy shows for what proportion of the refugee population living in each region age-disaggregated data was available at the end of 2021. While demographic coverage is close to universal for refugees hosted in Africa, it is available for `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Europe", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent of refugees in Europe, `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Asia", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent in Asia and only for `r round(as.numeric( t.missing.asyregion |> filter(asylum_region == "Americas", missingAge=="Age not missing") |>select(freq.totalEndYear) )*100)` per cent in the Americas. This is to a large extent a result of the differing population data sources in these regions: While the individual demographic details of refugees in many countries in Africa and MENA are recorded in UNHCR's own case registration system proGres, population data in other regions often comes from government offices with varying degrees of availability of demographic data.

@@

p.oriregion.missingAge2 <- ggplot(data = t.oriregion.missingAge %>%
                                   filter(origin_region!="Oceania", missingAge == "Age missing") ,
                                                 aes(x = origin_region, 
                                                     y = total)) +
  geom_bar(stat = "identity", colour = unhcrPaletteBlack[1], fill = unhcrPaletteWhite[1], alpha = 0,
           size = 1, linetype = "solid")  +
  ggtitle("Age data missing") +
 # labs(y = "Number of refugees")  +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 45),
        axis.text.y = element_blank(),
        plot.title = element_text(hjust = 0.5),
        axis.title.x =  element_blank(),
        axis.title.y = element_blank(),
        panel.grid.major.x = element_blank()) +
  scale_y_continuous(labels = label_number(scale_cut = cut_si("m")))

@@

 p.top4.children.ori.asy
```

@fig-childrenDist shows the proportion of children as a function of the distance between country of origin and country of asylum for the nine largest refugee producing countries. Dots for countries of asylum that neighbour the respective origin country are blue. The proportion of children by and large decreases for countries of asylum that are further away, and it is higher for neighbouring countries.

```{r, message=FALSE, warning=FALSE}
#| label: fig-childrenDist
#| fig-cap: Proportion of children among refugees for nine large refugee origins by distance to country of asylum (dot size is proportional to number of refugees in the country of origin/asylum pair)
#| fig-height: 11
#| fig-width: 8
#| out-width: "100%"
#| out-height: "100%"
#| warning: false
#| error: false
#| echo: false
#| message: false

p.children.logdist <- ggplot(data = dem_longMissing_plots,
       aes(x = logDistance, y = childrenProp*100, colour = neighbor)) +
  geom_point(shape = 19, aes(size = total, alpha = 0.3)) +  # Add points to the plot
  scale_color_manual(values = c(unhcrPaletteBlue[3], unhcrPaletteBlack[1])) +
  facet_wrap(~ origin_country, scales = "free_x") + # Group the data by origin_iso3
  scale_size(guide = "none") +
  scale_alpha(guide = "none") +
  labs(x = "Distance (log)", y="Percentage of children\nin country of asylum", colour = "Neighbouring\ncountry") +
#  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, angle = 0),
        axis.text.y = element_text(size = 13),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 10),
        panel.grid.major.x = element_blank()) +
    theme(legend.title = element_text(size = 10), legend.position = "bottom")

 p.children.logdist
```

# Model

Considering the findings of the descriptive analysis with considerable differences in the distribution of countries of origin and asylum between refugees with and without available data, and the explanatory power of origins and countries of asylum for the demographic characteristics of refugee populations, ignoring the missing data (i.e. treating it as if it was the same as the available data) risks introducing bias into regional and global estimates of the proportion of refugee children. In the following, we propose a model to predict counts for the age/sex brackets of those country pairs with missing demographic data (the *NAs* in @tbl-data), which in turn will allow us to estimate regional and global proportions of children. Summarising the descriptive analysis of missingness and available age data, there are three main conclusions to be considered in predicting the proportion of children for refugee populations without available age data:

1.  We cannot assume the global proportion of children among refugees for whom demographic data is available is the same as the proportion we would see if data was available for populations with currently missing data.\
2.  Knowing which country of origin refugees are from gives us some information regarding the proportion of children we would expect. For example, looking at @fig-childrenOriAsy we would generally assume that there is a higher proportion of children in a refugee population from South Sudan than in refugees from Venezuela.\
3.  For each origin population, knowing which region of asylum a population is from can help us refine the estimate for the proportion of children based on the country of origin. If we know a population is from Syria and they reside in a country of asylum in Europe, we would assume the proportion of children to be lower than if they were living in a country in Asia.

@@

Multilevel models (also called hierarchical, or sometimes random effect, models) are well suited to pool information across geographic levels as described above. We fit a multilevel model with varying nested intercepts at the levels of country of origin, region of asylum and country of asylum. Since the data is available as aggregated counts in ten age/sex brackets that sum up to the population total, we choose a multinomial distribution to model the underlying probabilities of a refugee falling in each of the age/sex brackets depending on their country of origin and country of asylum. When estimating these probabilities, the model treats populations from the same origin country as being more similar to each other than those from different origins, and for each origin considers those residing in countries in the same region of asylum as more similar than those in different regions of asylum. We fit the model in a Bayesian framework that allows us to incorporate uncertainty in parameters and in predictions from those parameters in the final predictions.

@@

The prior on the global intercepts with the $log(\frac{\Pi_{k}}{\Pi_{4}})$ mean aims to provide sensible levels for the age/sex bracket proportions. It incorporates our prior belief that, for example, it is unlikely that all refugees in a country are girls aged 0 to 4, but otherwise is not overly narrow. The student-t priors on the group-level standard deviations of the intercepts of country of origin and region of asylum are not informative and only serve to regularise the parameter space. The prior on country of origin is slightly wider to allow less pooling (i.e. closer to a separate model for each country of origin) and thereby take into consideration the general idea of the model that age/sex distributions are idiosyncratic to populations of an origin country. The very narrow $N(0,0.01)$ prior on standard deviation of the country of asylum term $b_{k,a}$ merits further discussion: Including the term might seem unusual to start with since in most cases there is only one observation in the form of age/sex counts and total (i.e. one row in the dataset) per country of origin/asylum to start with (exception are countries in Latin America and the Caribbean with two populations displaced from Venezuela, refugees and others in need of international protection). On one hand, leaving out this parameter leads to overconfidence in predicting counts in country pairs with missing data since in that case the counts are drawn from the varying intercepts of the regions (which are observed in the data in most country of origin/region of asylum pairs), ignoring the uncertainty that comes with predicting a whole new country pair not used in fitting the model. On the other hand, allowing a wider standard deviation leads to unreasonably wide prediction intervals (sometimes including counts of zero children in populations of several thousand refugees). Restricting the prior by way of a very small standard deviation helps us constrain the predictive intervals to sensible ranges. We include a fixed effect for neighbouring country pairs since @fig-childrenDist shows that we can on average expect higher proportions of children among refugee populations in neighbouring countries of asylum.

To predict the count vector of age/sex brackets $\tilde{y_{i^{*}}}$ for a country pair $i^{*}$ without reported data and account for uncertainty of parameters $\theta$ and prediction, we draw for each country pair with missing data $m$ values from the posterior predictive distributions: $$p(\tilde{y_{i^{*}}} \mid y)
= 
\int p(\tilde{y_{i^{*}}} \mid \theta) \cdot p(\theta \mid y)
\, \textrm{d}\theta$$ {#eq-postpred}

@@

The model was fit with R-package brms. Reproducible code is available [here](https://github.com/sebastiansteinmueller/ref-demographics-imputation/){.external target="_blank"}. Annex I contains the model summaries.

# Results

```{r}
##### predictions #####

@@

### by region of asylum

@@

## Venezuela 
imputations.ven <- imputations %>% 
  filter(origin_iso3 == "VEN")

### world

@@

## Global and regional estimates

```{r, message=FALSE, warning=FALSE}
#| label: tbl-estimates
#| tbl-cap: Regional and global model estimates for the proportion of children
#| warning: false
#| error: false
#| echo: false
#| message: false

t.dem.asyreg.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  group_by(asylum_region) %>% 
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

t.dem.children <- dem_longMissing %>% 
  filter(!is.na(children)) %>%
  summarise(children = sum(children), total = sum(total)) %>% 
  mutate(childrenProp = children/total)

t.world.children <- tibble(
  asylum_region = "WORLD",
  childrenMeanProp = t.imp.children$childrenMeanProp,
  lowerUI = t.imp.children$lowerUI,
  upperUI = t.imp.children$upperUI,
  childrenProp = t.dem.children$childrenProp
)

t.worldregion.children <- bind_rows(t.imp.asyreg.children, t.world.children) %>%
  select(asylum_region, childrenProp, childrenMeanProp, lowerUI, upperUI) %>% 
  mutate(across(where(is.double),~.*100 ))

t.worldregion.children %>% 
  gt() %>%
  fmt_number(columns = 2:5, decimals = 1) %>%
  tab_options(table.font.size = "40%") %>%
  cols_label(asylum_region = md("Region<br>of asylum"),
         childrenProp=  md("Percentage of<br>children in<br>available data"),
        childrenMeanProp =md("Estimated<br>percentage with<br>imputed data" ) ,
       lowerUI  =md("Lower 95%<br>uncertainty") ,
        upperUI =md("Upper 95%<br>uncertainty"))

@tbl-estimates shows the percentage of children among refugees living in the five major world regions estimated with imputations from the model and compares them to the percentage we obtain using the available data only. The global estimate for the proportion of children drops by almost 5 per cent, from 47 per cent to 42 per cent, when using the imputed dataset. The 95% uncertainty interval for the global point estimate ranges from 41.7 to 42.9 per cent and does not include the available-data figure of 47 per cent. This indicates the standardisation by origins and regions of asylum through the model is relevant, and previously disseminated figures for the proportion of refugees who are children were most likely overestimates. 
Regional estimates from the imputed data are below the available-data figures with narrow uncertainty intervals for all regions apart from Oceania. The available-data figures are based on 1209 refugees with reported data. The model pools the low estimate from this small dataset towards the global means.

## Uncertainty checks for country level predictions

To assess whether the model estimates uncertainty sensibly, we can look at the range of predictions at country level. @fig-impChildSyr shows the distribution of the proportion of children over the 200 imputed datasets among Syrian refugees in four countries of asylum without reported data. The posterior predictive distributions are relatively wide, with the 95% posterior predictive interval for the proportion of children among Syrians in the USA ranging from 15 per cent to 43 per cent. This is a desirable property of the models since, in the absence of reported data on an origin population in a country of asylum, we would not be confident in overly narrow prediction intervals. For this reason, the model output should generally not be used on country level, and never to estimate point intervals on the level of a country of asylum.

```{r, fig.pos = "H"}
#| label: fig-impChildSyr
#| fig-cap: Uncertainty in posterior predictive distribution of the proportion of children among Syrian refugees in four countries of asylum
#| warning: false
#| error: false
#| echo: false
#| message: false

p.imputations.syr.children <- ggplot(data = imputations.syr %>% filter(asylum_iso3 %in% c("SWE", "USA",    "GBR", "GRC")),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

p.imputations.syr.children
```

@fig-impChildVen shows the same plot for Venezuelans in four South American countries. The distributions for the proportion of children among Venezuelans in Ecuador and Argentina are much narrower than for those in Brazil and Chile, or for Syrians in any of the countries shown above. In contrast to many other countries, demographic data is available in Argentina and Ecuador for some Venezuelan populations, leading to narrower prediction range in these countries.

```{r, fig.pos = "H"}
#| label: fig-impChildVen
#| fig-cap: Uncertainty in posterior predictive distribution of the proportion of children among Venezuelan refugees in four countries of asylum
#| warning: false
#| error: false
#| echo: false
#| message: false

p.imputations.ven.children <- ggplot(data = imputations.ven %>% filter(asylum_iso3 %in% c("ARG", "ECU", "CHL", "BRA"), popType=="VDA"),
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_slab() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 11, angle = 0),
        axis.text.y = element_text(size = 11),
        axis.title.x = element_text(size = 11),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank())

p.imputations.ven.children
```

@fig-impChildSyrReg shows the distribution of reported (point only) and predicted (intervals) proportions of children among Syrian refugees in all countries by world region. Demographic data is reported from all countries in Africa and there is consequently no uncertainty in data from these countries. In Oceania in contrast, no data was reported at all, leading to very large uncertainty intervals for Australia and New Zealand.

```{r, fig.pos = "H"}
#| label: fig-impChildSyrReg
#| fig-cap: Uncertainty in posterior predictive distribution of the proportion of children among Syrian refugees across regions
#| fig-height: 13
#| fig-width: 9
#| warning: false
#| error: false
#| echo: false
#| message: false
p.imputations.syr.reg.children <- ggplot(data = imputations.syr,
      aes(y=reorder(asylum_countryShort, childrenProp), x=childrenProp*100)) +
     stat_pointinterval() +
  labs(x = "Proportion of children") +
  theme_minimal() +
    theme(axis.text.x = element_text(size = 13, angle = 0),
        axis.text.y = element_blank(),
        axis.title.x = element_text(size = 16),
        axis.title.y = element_blank(),
        panel.grid.major.y = element_blank()) +
    facet_wrap(~ asylum_region, scales = "free", ncol = 2) +
  theme(strip.text = element_text(size = 16))

p.imputations.syr.reg.children

# Limitations

While the imputed dataset is useful in estimating regional and global estimates of the age/sex distribution of refugees, it cannot substitute data collected and compiled for and at the national level. Estimates from the imputed dataset should always be disseminated and communicated with uncertainty intervals, and using the imputed data to derive country-level estimates is not recommended. The modelling approach is top-down and simply uses geographic proximity as a means to improve available-data figures. Promising improvements would be to use multi-year data and incorporate components of population change to obtain more reliable estimates.

# Annex I: Model output
```{r}
#| label: tbl-model_fixed
#| tbl-cap: Model - fixed effects
#| warning: false
#| error: false
#| echo: false
#| message: false
m.ageonly.sum <- summary(m.ageonly)
m.ageonly.ranefsum <- t(m.ageonly.sum$random %>% as_tibble() )
colnames(m.ageonly.ranefsum) <- c("female_0_4", "female_5_11" , "female_12_17" , "female_60" ,
                                  "male_0_4"  ,               "male_5_11"   ,            
"male_12_17"     ,          "male_18_59"     ,          "male_60")

m.ageonly.ranefsum <- m.ageonly.ranefsum %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(`Group and summary`=rowname)

m.ageonly.fixedsum <- m.ageonly.sum$fixed

m.ageonly.fixedsum %>% gt()
